Malacology: A Programmable Storage System Built on Ceph

### 2016-05-10 paper/paper.tex
Merge section 3 into 2
Also removes commented out text to avoid confusion

\if{false}
For example, a Ceph cluster contains a small number of
monitoring nodes (MONs) that use PAXOS to maintain a consistent view of
system-wide metadata such as data distribution parameters and cluster
membership. The data distribution metadata maintained by
the cluster monitors is used by RADOS servers and clients to provide
consistent reads and writes, and both RADOS and the monitoring services
are used by a cluster of file system metadata servers that provide POSIX
semantics using sophisticated indexing and distributed locking services.
We contend that re-using and re-purposing these code-hardened subsystems
is paramount to successfully adapting storage systems to new APIs and
new storage devices without losing the benefits from years of
code-hardening work.
\fi

@@

@@

\paragraph*{Malacology:} Map modification for propogation of state. We allow applications to augment
these data structures with application-specific metadata. In order to keep the
size of these data structures small application state may be stored elsewhere
and the references stored within the data structure.

@@

\subsection{Metadata Load Balancing}

@@

\section{Malacology Implementation}
\label{implementation}

In this section we describe how Malacology builds upon existing
services within Ceph. The main idea behind Malacology is to take the subsystems 
described in the previous section and expose APIs on top of them that allow 
users to dynamically inject code, allowing them to customize the behavior of 
these components. Supporting dynamic reconfiguration and programmability 
efficiently is critical in real systems that must adapt to a changing 
environment where potentially
conflicting policies arise. We've adopted Lua as the dynamic runtime  to
implement dynamic components in Malacology.

There are multiple reasons that make Lua an attractive runtime for implementing 
Malacology. Lua is a portable, embedded
scripting language that offers superious performance and productivity
trade-offs, including a JIT-based implementation that is well known for near
native performance. Additionally, Lua has been used extensively in game engines, 
and systems research \cite{neto:dls14-luaos}, including storage systems where it 
has been effectively used both on
\cite{grawinkel:pdsw2012-lua,watkins2013:bdmc13-in-vivo,geambasu_comet_2010} and 
off
\cite{sevilla:sc15-mantle} the performance critical path. Finally, the 
flexibility of the runtime allows to sandbox applications in order to address 
security and performance concerns. We now describe how Malacology exposes each 
subcomponent in Ceph.

\subsection{Object Storage}

% todo:
%  - ellude to the need for interfaces to be avialable quickly to support
%    our experiment about propogtation delay
%

The implementation of Ceph's object abstraction, although powerful, has some 
drawbacks.
Supporting only C/C++ for object interface developers, Ceph requires
distribution of compiled binaries for the correct architecture, adding a
large barrier of entry for developers and system administrators. Second,
having no way to dynamically unload modules, any changes require a full
restart of an OSD which may have serious performance impacts due to a loss
of cached data. And finally, the security limitations of the framework
limit the use of object classes to all but those with administrative level
access and technique expertise.

\subsection{Metadata Services}

% todo:
%  - ellude to the need for policies to be avialable quickly to support
%    our experiment about propogtation delay
% Original: inject strings
% New: use C++ bindings

Malacogy adds interface hooks in strategic spots in the MDS to control the load balancing and capability subsystems.

\subsubsection{Balancing Policy}

@@

\subsubsection{Capabilities}

\subsection{Cluster State}

Map modification for propogation of state. We allow applications to augment
these data structures with application-specific metdata. In order to keep the
size of these data structures small application state may be stored elsewhere
and the references stored within the data structure.

\subsection{Programmability and Durability}

{\bf Durability.}

Metadata service policies and object storage interfaces are stored durability
within RADOS and managed by storing references with the object maps. Since
the cluster already propagates a consistent view of these data structures,
we use this service to automatically install interfaces in OSDs, and install
policies within the MDS such that clients and daemons are synchronzied on
correct implementations without restarting.

\iffalse
\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  additional interface hooks and classes for OSDs and MDSs using the
  existing Ceph \texttt{cls} infrastructure
\item
  durable interface classes using the Ceph object store
\item
  versioned and consistent interfaces using the Ceph MONs
\end{enumerate}

Figure \ref{implementation} shows how interfaces can be loaded into
either the MDS or OSD (1), how interfaces are persisted to object
storage (2), and how they are versioned by the MONs (3). In the
following section, we will discuss the existing Ceph infrastructure and
the changes necessary for Malacology. We emphasize here that these
implementations are not necessarily optimized but they help demonstrate
the power of programmability.

\subsection{Additional Interface
Frameworks}\label{additional-interface-frameworks}

For object interfaces, clients send Lua classes to the OSD and then they
can invoke operations in that class remotely on the OSD. In reality,
this feature is a statically loaded object class written in C/C++ that
runs dynamically defined object interfaces, where clients access the
classes using an \texttt{exec()} wrapper. Although sending the class
with each client request is stateless (which has its own set of nice
features), it is costly for network bandwidth and difficult to organize
at the application level. Also, while the execute wrapper simplifies the
implementation, it burdens the applications, since they need to be
recompiled to use the \texttt{exec()} function.

Although these OSD interfaces are powerful, the current implementation
has drawbacks. OSD interfaces are written in C/C++ and compiled into a
shared library, so the developer must account for different target
architectures. Second, C/C++ provides more functionality than is
necessary since the snippets of interface code mainly set policies and
perform simple operations. Finally, developing these interfaces in C/C++
has high overhead. The developer must learn how the OSD dynamically
loads the shared libraries (\emph{e.g.}, OSDs rely on a strict naming
convention to find shared libraries), how to get their interfaces
compiled using the Ceph \texttt{make} system, and how to debug issues
that are not related to their specific interface (\emph{e.g.}, the OSD
cannot find the shared library) - this learning curve is unacceptable
for non-Ceph developers, especially since most interfaces are one-off
solutions specific to their applications.

\subsubsection{New Interface Classes}\label{new-interface-classes}

Our framework has added a mechanism for defining and running object and
metadata balancer classes using the Lua scripting language. Scripting is
useful in large systems for 3 reasons:

\begin{itemize}
\item
  accelerates deployment: tweaking policies, like changing a threshold
  or metric calculation, does not require a full recompile or system
  start-up.
\item
  transparency: if the bindings carefully expose the correct metrics and
  functions, the user does not need to learn the internal functions,
  variables, or types of the system.
\item
  portability: scripts can be exchanged across platforms and system
  versions without compatibility issues.
\end{itemize}

Our Lua bindings expose functions and symbols both ways; the host
program can call functions defined in Lua and the Lua scripts can call
functions defined in native C++. These bindings are merged upstream. Lua
is frequently used in game engines to set policies but we use it here
because most of the user-defined classes in Ceph are policies as well!
We do not want to provide specific implementations, like pulling data
from objects or transferring them over the network, but instead strive
to say \emph{what to do} with the data once we have it. Separating
policy from mechanism is a driving factor in using Lua. We choose Lua
for 4 reasons: performance, portabability, size, and security.

Lua performs well as an embeddable scripting language. Although it is an
interpreted language, the bindings are designed to make the exchange of
variables and functions amongst languages straightforward. The LuaJIT
virtual machine boasts near-native performance, making it a logical
choice for scriptability frameworks in systems research
\cite{neto:dls14-luaos}. In storage systems, it has been effectively
used both on \cite{grawinkel:pdsw2012-lua,watkins2013:bdmc13-in-vivo} and off
\cite{sevilla:sc15-mantle} the critical path, where performance is
important.

Because Lua is interpreted, it is also portable. The portability allows
functions to be shipped around the cluster without having to compile or
reason about compatibility issues. Traditionally, intefaces are written
in C/C++, so they must be compiled into shared libraries before being
injected into the daemon at runtime. Shared libraries need to be
compiled on the host they will be run on to accomodate different
architectures and dependendencies. Embedding the Lua interpretator, on
the other hand, gives us the flexbility to store the actual Lua code in
other places, like RADOS, so we can enjoy all the other properties
provided by that particular back-end.

The other two advantages, that Lua leaves a small memory footprint and
that it is secure, are discussed at length in 
\cite{ierusalimschy_programming_2006,neto:dls14-luaos}. In short, Lua
provides relatively robust sandbox capabilities for protecting the
system against some malicious cases, but does not protect the system
from poor policies. Since Malacology does not use these features (yet),
we will avoid discussion here.

\subsection{New Interface Hooks}\label{new-interface-hooks}

\fi

\iffalse

\subsection{Programmability and Durability}

Traditionally, object storage interfaces are compiled into shared
libraries and loaded into a running OSD daemon using \texttt{dlopen()}.
Because the shared library has to link against symbols found in the
running executable, the interfaces are stored on the local file systems
of each OSD so they can be versioned and distributed alongside the same
Ceph binaries. This policy is for safety, since the OSDs could be on
different distros that provide different versioning capabilities. Also,
this approach allows bug fixes to co-evolve with the rest of the Ceph
installation.

Because of the portability of Lua and because the Lua virtual machine is
compiled and loaded as a shared library, the MDS/OSD can store the
balancers in a more redundant location. For our implementation, we store
balancers in RADOS and the daemons first search RADOS, then the local
file system for their interface classes.

\subsection{Versioning Interfaces with
MONs}\label{versioning-interfaces-with-mons}

We added two new commands to Ceph command line client so the user can
inject interface versions into the cluster::

\noindent \texttt{ceph\ osd\ pool\ set-class\ \textless{}pool\textgreater{}\ \textless{}class\textgreater{}\ \textless{}script\textgreater{}}

\noindent \texttt{ceph\ mds\ set\ lua\_balancer\_class\ \textless{}version\textgreater{}}

By leveraging the monitor daemons, our services get the consistency from
the monitors version management, the distribution from the monitors data
structures (which are already distributed), and the durability from the
robustness of the monitors and persistence with Paxos. The
implementation uses the placement group pool data structure which
maintains the policies (\emph{e.g.}, erasure coding) and organization
details (\emph{e.g.}, snapshots) for each pool. While stuffing the
information into the monitor map, the structure that describes the the
monitor topology, was an option, the placement group pool allows finer
grain control over different types of data.

In regards to the consistency, each time the user injects a new Lua
object class it enters the monitor quorum as a proposal; updates to the
placement group pool need to wait until accepted, at which point the
state will be propogated and versioned. We hooked up the monitor command
to the \texttt{cls\_lua} class by having the OSD pull the script from
the placement group instead of trying to dynamically load the shared
library with \texttt{dlopen()} or the Lua script.

The cls-lua branch lets users write object classes in Lua, the lua-rados
client library lets users write applications that talk to RADOS with
Lua, and the cls-client uses the lua-rados library to send Lua object
classes to the OSDs equipped with the LuaJIT VM.
\fi

>>>>

@@

\paragraph*{Malacology:} Metadata service policies and object storage interfaces are stored durability
within RADOS and managed by storing references with the object maps. Since
the cluster already propagates a consistent view of these data structures,
we use this service to automatically install interfaces in OSDs, and install
policies within the MDS such that clients and daemons are synchronzied on
correct implementations without restarting.

@@

\paragraph*{Malacology:} Malacology exposes the monitoring service in two ways. 
First, Malacology APIs for object storage (\S\ref{active-storage}) and the MDS 
(\S\ref{malacology:mds}) have functions available to them that allow them to 
store application-specific data in the highly-available, consistent store 
associated to the Ceph monitors. As mentioned
Map modification for propogation of state. We allow applications to augment
these data structures with application-specific metadata. In order to keep the
size of these data structures small application state may be stored elsewhere
and the references stored within the data structure.

@@

\paragraph*{Malacology:} Malacogy adds interface hooks in strategic spots in the MDS to control the load balancing and capability subsystems.
Malacogy adds interface hooks in strategic spots in the MDS to control the load balancing and capability subsystems.

\subsection{Metadata Load Balancing}
\label{malacology:mds}

@@

@@
